Classless Java: Tuning Java Interfaces

### 2016-03-15 OOPSLA16/sections/Implementation.tex
Some more text. Table and flow chart.

@@

The annotation we created is \mixin. In Eclipse, with an interface annotated by
\mixin, the automatic annotation processing is performed transparently and the information of
the interface from compilation is captured in the ``Outline'' window. This includes
all the methods inside the interface as well as the generated ones.  The custom
transformation is easy and convenient to use.  For example, this means that the
IDE functionality for content assist and autocomplete will work for the newly generated
methods. The biggest reasons to use Lombok rather than using a conventional Java
 annotation processor are:
\begin{itemize}
%this seems like a commercial spot
%\item Lombok is byte-code based instead of source-code based, which makes client
%  code concise and easy to maintain. Such code generation is performed at
%  compile time to modify bytecode.
\item Lombok modifies the generation process of the class files, by directly modifying the AST.
Neither the source code is modified nor new Java files are generated.
\item Moreover, and probably more importantly, Lombok is capable of generating
  code \emph{inside} a class/interface, which conventional
  Java annotation processors do not support.
\end{itemize}

\noindent 
Lombok offers several advantages over conventional code preprocessors/macros, 
and our disciplined \mixin annotation offer some more advantages with respect to arbitrary Lombok based
ast rewriting.
modularity, code quality and maintainability
\begin{itemize}
\item \textbf{Modularity}
While general preprocessing can act across module boundaries, Lombock act modularly on each class, and it makes
sense to apply the transformation to one class/interface at a time, and only to annotated classes/interface.
This allows library code to be reused without the need of being
reprocessed and recompiled.

\item \textbf{Syntax and type errors.} 
Some preprocessors (like the C one) can produce syntactically invalid code.
Lombok ensures only syntactically valid code is produced; however, type errors can appear.
In our annotation, we proven that we do not introduce new errors in clients (see [] for details).

\item \textbf{Obfuscated source code.}
Preprocessors bring great power, that can easily be misused producing 
code particularly hard to understand.
Lombok requires to start from Java Syntax, and can re interpret it.
This is already better, since the syntax is preserved, but the reinterpretation could be surprising for 
badly designed Lombok based rewritings.
However, the``Outline'' window of Eclipse or other IDEs (and thus their auto-completion facilities)
will show the result of the rewriting, helping the programmer to use the reinterpreted semantic.

>>>>

\begin{figure*}[t]\label{fig:lombok}
\centering
\includegraphics[width=5in]{pdfs/lombok.png}
\caption{The flow chart of Lombok annotation processing. [Ref: http://notatube.blogspot.hk/2010/12/project-lombok-creating-custom.html]}
\end{figure*}

@@

The annotation we created is \mixin. Figure~\ref{fig:lombok}\haoyuan{Wrong ref?} illustrates the
flow of \mixin annotation processing in Lombok. In Eclipse, with an interface annotated by
\mixin, the automatic annotation processing is performed transparently and the information of
the interface from compilation is captured in the ``Outline'' window. This includes
all the methods inside the interface as well as the generated ones.  The custom
transformation is easy and convenient to use.  For example, this means that the
IDE functionality for content assist and autocomplete will work for the newly generated
methods. The biggest reasons to use Lombok rather than using a conventional Java
 annotation processor are:
\begin{itemize}
%this seems like a commercial spot
%\item Lombok is byte-code based instead of source-code based, which makes client
%  code concise and easy to maintain. Such code generation is performed at
%  compile time to modify bytecode.
\item Lombok modifies the generation process of the class files, by directly modifying the AST.
Neither the source code is modified nor new Java files are generated.
\item Moreover, and probably more importantly, Lombok is capable of generating
  code \emph{inside} a class/interface, which conventional
  Java annotation processors do not support.
\end{itemize}

\noindent
Lombok offers several advantages over conventional code preprocessors/macros,
and our disciplined \mixin annotation offer some more advantages with respect to arbitrary Lombok-based
AST rewriting.
\begin{itemize}
\item \textbf{Lack of modularity.}
While general preprocessing can act across module boundaries, Lombok acts modularly on each class, and it makes
sense to apply the transformation to one class/interface at a time, and only to annotated classes/interfaces.
This allows library code to be reused without the need of being
reprocessed and recompiled.

\item \textbf{Syntax and type errors.}
Some preprocessors (like the C one) can produce syntactically invalid code.
Lombok ensures only syntactically valid code is produced; however, type errors can appear.
In our annotation, we prove that we do not introduce new errors in clients (see Section~\ref{subsec:guarantees} and Section~\ref{subsec:results} for details).

\item \textbf{Obfuscated source code.}
Preprocessors bring great power, which can easily be misused producing
code particularly hard to understand. Thus code quality and maintainability are reduced.
Lombok requires to start from Java syntax, and can reinterpret it.
This is already better, since the syntax is preserved, but the reinterpretation could be surprising for
badly designed Lombok-based rewritings.
However, the``Outline'' window of Eclipse or other IDEs (and thus their auto-completion facilities)
will show the result of the rewriting, helping the programmer to use the reinterpreted semantic.

\item \textbf{No reuse of the type system.}
As we mentioned above, badly designed rewritings can arise from the great power of Lombok. A simple piece of source code
\begin{lstlisting}
interface M { int m(); }
\end{lstlisting}
can be reinterpreted as
\begin{lstlisting}
interface M { void m(String s); }
\end{lstlisting}
in which case the type of method \Q@m@ is changed. Our \mixin annotation does not introduce this kind of rewritings,
and hence the type system is reused. Moreover, Lombok can also modify unbounded types, which is easy to understand,
for instance, the following code
\begin{lstlisting}
interface M { T m(); } // T is unbounded
\end{lstlisting}
is transformed into
\begin{lstlisting}
interface M { int m(); } // No error message
\end{lstlisting}
in which case the user will see the unbounded type in source code, but without error message from the compilation, since
Lombok has modified the return type of \Q@m@. However, our \mixin annotation can still keep such errors and warnings.

### 2016-03-15 OOPSLA16/sections/Overview.tex
Some more text. Table and flow chart.

\paragraph{Summary} Dealing with state (mutable or not) in object interfaces
relies on a notion of abstract state, where only \emph{methods} that interact with
state are available to users. Object interfaces provide support for four
different types of abstract state operations, which are summarized in
Figure~\ref{fig:abstractstate}.  The abstract state operations are determined by
naming conventions and the types of the methods. Fluent setters are a variant of
conventional setters, and are discussed in more detail in Section~\ref{sec:dsls}.

\begin{figure*}
\centering
\saveSpaceFig
\begin{tabular}{|l|l|l|}
\hline
 & \textbf{Example}                  & \textbf{Description } \\ \hline
\textbf{``fields''/getters}        &   \Q@int x();@                  & Retrieves value from field \texttt{x}.          \\ \hline
{\bf withers}        &   \Q@Point2D withX(int val);@                & Clones
object; updates field \texttt{x} to \texttt{val}.             \\ \hline
\textbf{setters}        & \Q@void x(int val);@ & Sets the field
\texttt{x} to a  new value \texttt{val}.        \\ \hline
\textbf{fluent setters}        & \Q@Point2D x(int val);@ &Sets the field
\texttt{x} to \texttt{val} and returns \texttt{\this}.           \\ \hline
\end{tabular}

\caption{Abstract state operations, for a field \texttt{x}, allowed by the \mixin
  annotation. \marco{I disagree with this table, for example we miss the with} }

>>>>

\paragraph{Summary} Dealing with state (mutable or not) in object interfaces
relies on a notion of abstract state, where only \emph{methods} that interact with
state are available to users. Object interfaces provide support for four
different types of abstract state operations \textcolor{red}{together with two types of
general operations}, which are summarized in
Figure~\ref{fig:abstractstate}.  The abstract state operations are determined by
naming conventions and the types of the methods. Fluent setters are a variant of
conventional setters, and are discussed in more detail in Section~\ref{sec:dsls}.

\begin{figure*}
\centering
\saveSpaceFig
\begin{tabular}{|l|l|l|l|}
\hline
& \textbf{Operation}  & \textbf{Example}                  & \textbf{Description } \\ \hline
\multirow{4}{*}{State operations} & \textbf{``fields''/getters}        &   \Q@int x();@                  & Retrieves value from field \texttt{x}.          \\ \cline{2-4}
& {\bf withers}        &   \Q@Point2D withX(int val);@                & Clones
object; updates field \texttt{x} to \texttt{val}.             \\ \cline{2-4}
& \textbf{setters}        & \Q@void x(int val);@ & Sets the field
\texttt{x} to a  new value \texttt{val}.        \\ \cline{2-4}
& \textbf{fluent setters}        & \Q@Point2D x(int val);@ &Sets the field
\texttt{x} to \texttt{val} and returns \texttt{\this}.           \\ \hline
\multirow{2}{*}{Other operations} & \textbf{factory methods} & \begin{tabular}{l}\Q@static Point2D of(int _x,@\\ \hspace{.8in}\Q@int _y) {...}@\end{tabular} & Generate factory method for \texttt{Point2D}.        \\ \cline{2-4}
& \textbf{general updaters}        & \Q@Point3D with(Point2D val);@ & Update all fields
that match the object \texttt{val}.        \\ \hline
\end{tabular}

\caption{\textcolor{red}{(edited)} Abstract state operations for a field \texttt{x}, together with general operations, supported by the \mixin
  annotation. \marco{I disagree with this table, for example we miss the with} }
